A Global Optimization Algorithm for Non-Convex Mixed-Integer Problems

### 2018-01-11 2018/LION_Mixed-Integer/LION2018.tex
After translation

\usepackage{amsmath}
\usepackage{enumitem}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}
\usepackage[T2A]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[russian,english]{babel}

@@

\title{A Global Optimization Algorithm for Non-Convex Constrained Mixed-Integer Problems}
\author{Victor Gergel \and Konstantin Barkalov \and Ilya Lebedev %\Letter 
\\
\email{konstantin.barkalov@itmm.unn.ru}}

@@

\begin{abstract}
\Russian
В работе рассматриваются mixed-integer global optimization problems. Предложен новый детерминированный алгоритм для решения задач данного класса, основанный на information-statistical approach к решению задач непрерывной глобальной оптимизации. Проведено сравнение данного алгоритма с известными аналогами, показывающее эффективность разработанного подхода. Устойчивая работа алгоритма подтверждена также решением серии из нескольких сотен mixed-integer global optimization problems.

@@

\section{Introduction}\label{sec:intro}
\Russian
В работе рассматриваются global optimization problems и методы их решения. Задачи глобальной оптимизации обладают значительной вычислительной трудоемкостью, т.к. глобальный оптимум является интегральной характеристикой решаемой задачи и требует исследования всей области поиска. Как результат, поиск глобального оптимума сводится к построению некоторого покрытия (вообще говоря, неравномерного) в области параметров. Особый интерес представляют задачи, в которых часть параметров может принимать лишь целочисленные значения (mixed-integer global optimization problems), т.к. для них сложнее построить оценки оптимума по сравнению с непрерывными задачами.

Методам решения mixed-integer problems посвящена обширная литература (см., например, обзоры \cite{Burer,Boukouvala}). Известные детерминированные методы решения задач данного класса основаны, как правило, на подходах Branch-and-Bound or Branch-and-Reduce. Также известен ряд метаэвристических и генетических алгоритмов, так или иначе основанных на идеях случайного поиска.

В данной работе нами предложен новый детерминированный метод решения mixed-integer problems, основанный на information-statistical approach к решению задач глобальной оптимизации \cite{Strongin2000,Strongin2013}. Текст статьи, отражающей предварительные результаты исследования, построен следующим образом. Вначале рассмотрен подход к решению задач с непрерывными параметрами, затем предложено его обобщение для mixed-integer problems. В финальном разделе приведены результаты сравнения предложенного метода с известными аналогами.

@@

Let us consider a multiextremal optimization problem in the form
\begin{gather}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.\nonumber
\end{gather}
Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints satisfy the Lipschitz condition
\[ 
\left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, 
\]
with constants $L_i, \; 1 \leq i \leq m+1$, respectively,  and may be multiextremal.

@@

%Проверить
An efficient global search algorithm (GSA) for solving constrained optimization problem (\ref{problem1}) was developed at University of Nizhni Novgorod. A detailed description of the algorithm and the corresponding convergence theory are presented in \cite{Strongin2000}.

@@

\Russian
Рассмотрим теперь способ адаптации GSA для решения mixed-integer global optimization problems

\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.\nonumber
\end{gather}

На основе задачи (\ref{problem_i}) сформируем множество задач 
\begin{gather}\label{problem_is}
\min{\left\{ g_{m+1}^s(y):y\in D^s, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\},\nonumber 
\end{gather}
каждая из которых соответствует исходной задаче (\ref{problem_i}) с фиксированным набором целочисленных параметров. Число задач $S$ будет соответствовать числу возможных комбинаций целочисленных параметров.

Используя схему редукции размерности с помощью развертки $y(x)$ на основе множества задач (\ref{problem_is}) можно составить единую задачу
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}^s(y(x)): x \in [0,S], \; g_i^s(y^s(x)) \leq 0, \; 1 \leq i \leq m\right\},
\end{equation}
где отображение $y^s(x)$  формируется на основе исходной развертки $y(x)$ следующим образом
\[
y^s(x)=y(x-s), \; x\in[s-1,s],\; s\in\{1,...,S\}.
\]
Функции $g_i^s(y^s(x))$ могут иметь разрывы в точках $s\in \{1,...,S-1\}$, поэтому данные точки рассматриваются как выколотые, значение objective function and constraints в них не определено.

Для иллюстрации на Fig. \ref{fig:1}  изображены графики функций, соответствующих задаче с одним непрерывным и одним бинарным параметром.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.7\textwidth]{fig1.jpg}
    \caption{Mixed-integer global optimization problem}
    \label{fig:1}
\end{figure}

Применяя к решению задачи (\ref{problem_is1}) алгоритм глобального поиска, мы найдем решение задачи (\ref{problem_i}). При этом основная часть испытаний будет проведена в $s$-й подзадаче, решение которой соответствует решению исходной задачи (\ref{problem_i}). В остальных подзадачах будет проведена лишь незначительная часть испытаний, т.к. решения данных подзадач являются локально-оптимальными по отношению к решению $s$-й подзадачи. Сказанное подтверждается Fig. \ref{fig:1}, где штрихами обозначены точки испытаний, выполненных при решении данной задачи.

Таким образом, мы сформировали Mixed-Integer Global Search Algorithm (MIGSA), основанный на сведении задачи MINLP путем к задаче NLP. Можно доказать, что условия сходимости данного алгоритма будут следовать из условий сходимости его прототипа (GSA).

@@

\Russian
Проведем сравнение предложенного метода MIGSA с genetic algorithm for solving mixed-integer optimization problems, реализованным в Matlab Global Optimization Toolbox. В Table \ref{tab:1} приведено число испытаний, потребовавшееся для решения известных тестовых mixed-integer problems данными методами. Для обоих методов была использована одинаковая точность поиска решения $10^{-2}$. Все вычислительные эксперименты были проведены на компьютере с процессором Intel Core i5-7300 2.5 GHz и 8 Gb RAM под управлением MS Windows 10. Результаты экспериментов показывают превосходство метода MIGSA как по числу итераций, так и по времени работы.

\begin{table}
	\caption{Сравнение эффективности методов MIGSA и GA}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & $k$ & $t$ & & $k$ & $t$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & &	417 &	0.04 \\
%		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & &	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & &	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & &	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 & & 372	 &	0.017 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

Для демонстрации надежности метода MIGSA нами были решены четыре серии по 100 многоэкстремальных mixed-integer problems, построенных на основе модифицированных классов задач Simple и Hard, сгенерированных генератором GKLS \cite{Gaviano}. 
Данный генератор многоэкстремальных функций часто используется для исследования алгоритмов непрерывной оптимизации \cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
 В сгенерированных нами задачах было 5 целочисленных и 3 непрерывных параметра. Точность поиска решения равнялась $10^{-2}$ по координате. Все задачи из серии были успешно решены , при этом для решения задач на основе класса Simple потребовалось среднем 11988 trials, а для решения задач на основе класса Hard -- 24750 trials.

@@

\bibitem{Boukouvala}
Boukouvala, F., Misener, R., Floudas, C.A.: Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO. European Journal of Operational Research \textbf{252}, 701--727 (2016)

@@

\bibitem{Gaviano}
Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)

>>>>

\usepackage{amsmath}
\usepackage{enumitem}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{epsfig}

@@

\title{A Global Optimization Algorithm for Non-Convex Mixed-Integer Problems}
\author{Victor Gergel \and Konstantin Barkalov \and Ilya Lebedev %\Letter 
\\
\email{konstantin.barkalov@itmm.unn.ru}}

@@

\begin{abstract}
In the present paper, the mixed-integer global optimization problems are considered. A novel 
deterministic algorithm for solving the problems of this class based on the information-statistical 
approach to solving the continuous global optimization problems has been proposed. The 
comparison of this algorithm with well known analogs demonstrating the efficiency of the 
developed approach has been conducted. The stable operation of the algorithm was confirmed 
also by solving a series of several hundred mixed-integer global optimization problems.

@@

\section{Introduction}\label{sec:intro}

In the present paper, the global optimization problems and the method of solving these ones are 
considered. The global optimization problems are the computation-costly ones since the global 
optimum is an integral characteristic of the problem being solved and requires the investigation 
of the whole search domain. 
%As a result, the search of the global optimum is reduced to the construction of a coverage (in general, a nonuniform one) of the space of parameters. 
The problems, in which some parameters can take the inter values only (mixed-integer global 
optimization problems) are of special interest because for these ones it is more difficult to build 
the estimates of the optimum as compared to the continuous problems.

A lot of publication have been devoted to the methods of solving the mixed-integer problems 
(see, for example, the reviews~\cite{Burer,Boukouvala}). The well known deterministic 
methods of solving the problems of this class are based, as a rule, on the Branch-and-Bound 
approach or on the Branch-and-Reduce one. Also, a number of the metaheuristic and genetic 
algorithms are known based 
ones, anyway, based on the ideas of random search.

In the present study, we proposed a novel deterministic method for solving the mixed-integer 
problems based on the information-statistical approach to solving the global optimization 
problems \cite{Strongin2000,Strongin2013}. The paper text reflecting the results of preliminary 
studies is composed as follows. First, an approach to solving the problems with continuous 
parameters has been considered, then its generalization for the mixed-integer problems was 
proposed. In the final section, the results of comparison of the proposed method with the well 
known analogs are presented.

@@

Let us consider a multiextremal optimization problem in the form
\begin{gather}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.\nonumber
\end{gather}
Suppose, that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints satisfy the Lipschitz condition
%\[ \left|g_i(y')-g_i (y'')\right| \leq L_i \left\|y'-y'' \right\|, \; y',y''\in D, \; 1\leq i \leq m+1, \]
%with constants $L_i, \; 1 \leq i \leq m+1$, respectively,  
and may be multiextremal.

@@

An efficient global search algorithm (GSA) for solving the constrained optimization problem 
(\ref{problem1}) has been developed at University of Nizhni Novgorod. A detailed description 
of the algorithm and the corresponding convergence theory are presented in 
\cite{Strongin2000}.

@@

Now, let us consider a method of adaptation of GSA for solving the mixed-integer global 
optimization problems 
\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N, \; y_j \in Z, \; j \in J, \; y_i \in R, \; i \notin J \right\}.\nonumber
\end{gather}

Let us formulate a set of problems based on the problem (\ref{problem_i})
\begin{gather}\label{problem_is}
\min{\left\{ g_{m+1}^s(y):y\in D^s, \; g_i^s(y)\leq 0, \; 1 \leq i \leq m\right\}},\\
D^s=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, \; i \notin J \right\}, \; s\in\{1,...,S\}.\nonumber 
\end{gather}
Each of these problems corresponds to the initial problem (\ref{problem_i}) with fixed set of 
integer parameters. The number of problems $S$ will correspond to the number of possible 
combinations of the integer parameters.

Using the dimensionality reduction scheme utilizing the evolvent $y(x)$ 
%based on the set of problems (\ref{problem_is}),
one can compose a united problem 
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}^s(y^s(x)): x \in [0,S], \; g_i^s(y^s(x)) \leq 0, \; 1 \leq i \leq m\right\},
\end{equation}
where the mapping $y^s(x)$ is formed on the basis of initial evolvent $y(x)$ as follows
\[
y^s(x)=y(x-s), \; x\in[s-1,s],\; s\in\{1,...,S\}.
\]
The functions $g_i^s(y^s(x))$ can have the breaks at the points $s\in \{1,...,S-1\}$, therefore, 
these points were considered as the punctured ones, the values of the objective function and 
constraints at these points remain undefined.

As an illustration, in Fig. \ref{fig:1} presents the plots of the functions corresponding to a 
problem with one continuous parameter and one binary parameter.

\begin{figure}[ht]
    \centering
    \includegraphics[height=2.9cm,width=0.8\textwidth]{fig1.jpg}
    \caption{Mixed-integer global optimization problem}
    \label{fig:1}
\end{figure}

Applying the global search algorithm to solving the problem (\ref{problem_is1}), we will find 
the solution of the problem (\ref{problem_i}). In this case the major part of trials will conducted in 
the $s$-th subproblem, the solving of which corresponds to the solving of the initial 
problem (\ref{problem_i}). In the rest subproblems, a minor part of trials only will be performed 
since the solutions of these subproblems are the locally optimal ones with respect to the solution 
of the $s$-th subproblem. All the above is confirmed by Fig. \ref{fig:1}, where the 
points of trials executed in the course of solving this problem are denoted by the dashes.

Thus,  we have constructed the \textit{Mixed-Integer Global Search Algorithm} (MIGSA) based on the  
reduction of the MINLP to the NLP. 
%It is possible to prove that the convergence conditions for this algorithm will follow from the convergence conditions of its prototype (GSA).

@@

Let us compare the proposed MIGSA method with a genetic algorithm for solving the mixed-
integer optimization problems implemented in Matlab Global Optimization Toolbox. In Table 
\ref{tab:1}, the numbers of trials required for solving the well-known test mixed-integer 
problems by these methods are presented. For both methods, the same values of the accuracy of 
the solution search $10^{-2}$ were used. All numerical experiments were conducted on a 
computer with Intel Core i5-7300 2.5 GHz and 8 Gb RAM processor under MS Windows 10. The 
results of experiments have demonstrated the advantage to the MIGSA method in the number 
of iterations as well as in the execution time.

\begin{table}
	\caption{Comparison of MIGSA and GA}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
	\multirow{2}{*}{Test problem}	 & \multicolumn{2}{c}{ GA } & & \multicolumn{2}{c}{MIGSA} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & $k$ & $t$ & & $k$ & $t$  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
		 Problem 2 \cite{Floudas}&	481 &	0.0601 & &	417 &	0.04 \\
%		 Problem 3 \cite{Floudas}& 	1821 &	0.1130 & & 3324 &	0.107 \\
		 Problem 6 \cite{Floudas}&	641 &	0.0510 & &	118 &	0.001 \\
		 Problem 1 \cite{Deep}   &	481 &	0.1378 & &	66 &	0.0007 \\
		 Problem 2 \cite{Deep}   &	481 &	0.0473 & &	57 &	0.0006 \\
		 Problem 7 \cite{Deep}   &	841 &	0.0736 & & 372	 &	0.017 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 
multiextremal mixed-integer problems each constructed on the basis of modified problems of 
\textit{Simple} and \textit{Hard} classes generated by GKLS generator. %~\cite{Gaviano}. 
This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
In the problems generated in our experiments, there were 5 integer parameters and 3 continuous 
ones. The accuracy of the search %of the solution in the coordinates 
was equal to $10^{-2}$. All  
problems of the series have been solved successfully, at that, for solving the problems based on 
the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based 
on the \textit{Hard} class 24750 trials were required.

@@

\bibitem{Boukouvala}
Boukouvala, F., Misener, R., Floudas, C.A.: Global optimization advances in Mixed-Integer Nonlinear Programming, MINLP, and Constrained Derivative-Free Optimization, CDFO. European J. Oper. Res. \textbf{252}, 701--727 (2016)

@@

%\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)
